\huge A Usability Evaluation of Tor Launcher

### 2017-03-13 pets2017-paper/pets2017-paper.tex
fixed easy rewording suggestions from shepherd

\section{Introduction}
Tor~\cite{dingledine2004tor} is an anonymity network that routes Internet traffic through a series of relays 
that make it difficult to observe sources and destinations. Tor Browser~\cite{torbrowser} is a modified Firefox browser that routes traffic through Tor, thereby simplifying the process of configuring Tor to use existing browsers. Although Tor Browser was originally designed for anonymity, many now use Tor Browser to circumvent Internet censorship. Many countries block Tor relays specifically to prevent their citizens from circumventing censorship~\cite{winter2012great}. Tor Browser is bundled with a graphical configuration interface, Tor Launcher (Figure~\ref{fig:old-interface}), which allows users to bypass censorship by configuring various bridges, proxies, and transport protocols prior to connecting to Tor. Of course, these configuration options are only effective if users understand how to correctly use them. Thus, we performed a usability study wherein we simulated various censorship environments in order to examine the effectiveness of Tor Launcher and how it could be improved.

Improving Tor launcher has the obvious benefit of helping censored users connect to Tor, but this also benefits other Tor users: an increase in overall users increases the anonymity set~\cite{dingledine2006anonymity}). In this paper, we describe how we started our usability evaluation by inspecting the interface (Section~\ref{sec:inspection}) and observing participants interact with it (Section~\ref{sec:qualitative}). Based on the insights from our observations, we proposed some changes to the interface (Section~\ref{sec:design}) and
quantified how usable the interface was before and after changes (Section~\ref{sec:quantitative}). Finally, we provide recommendations based on what we learned (Section~\ref{sec:recommendations}), we address
limitations of our evaluation (Section~\ref{sec:limitations}), and frame our study in the context of existing work (Section~\ref{sec:related}). We hope that our usability evaluation serves as an informative case study and provides guidance on further improving Tor Launcher, as well as other privacy-enhancing technologies.

@@

Bridges are unlisted alternative entry nodes used to connect to Tor when Tor relays are censored.
Bridges also receive packets from the user and forward traffic to middle relays, but also may run a pluggable transport (referred to as ``transport'' for shorthand). Pluggable transports specify alternative protocols by which Tor Browser may connect to the Tor network, thereby disguising Tor traffic to make it harder to detect and block bridges. For convenience, bridges are grouped by which transport they run.

Most bridges can be blocked by blocking their static IP addresses. These include bridges that run``fte'' and ``fte-ipv6''~\cite{fte} pluggable transports, which disguise the Tor protocol as another protocol (such as HTTP); these also include the ``obfs3''~\cite{obfs3}, ``obfs4''~\cite{obfs4}, and ``scramblesuit''~\cite{scramblesuit} transports, which encrypt or alter the Tor protocol to appear as random noise. Other bridges are less susceptible to blocking because of how they work. Bridges that run the ``flashproxy''~\cite{flashproxy} and ``meek''~\cite{fifield2015blocking} transports route traffic through third party web browsers and content delivery networks. The bridges that are accessible in each country vary. Figure~\ref{fig:old-bridge-combobox} shows the bridge and transport options at the time of our study.

@@

\subsection{Connecting to Tor} 
There are many valid configuration settings that can be used to connect to the Tor network.
A user who does not need a bridge or proxy can still connect with a bridge, proxy, or both, provided that they are configured correctly (though adding unnecessary network hops or obfuscation protocols may slow the connection down).

@@

\subsection{Testing Pipeline} 
We performed a cognitive walkthrough to inspect the interface to anticipate potential problems and formally assess its pros and cons. We then conducted qualitative and quantitative usability tests. We believe both methods were necessary for a complete evaluation. Qualitative testing tells you {\it why or how} it is or is not usable, while quantitative testing tells you {\it how much} it is or is not usable. For gathering qualitative data, five users is considered to maximize the cost-benefit ratio~\cite{howmanyusers}, so we recruited five participants for each condition---multiple simulated censorship environments---in our qualitative experiment. Twenty users is considered to be the minimum for yielding statistically significant measurements~\cite{howmanyusers}, so we recruited twenty participants for each condition in our quantitative experiment.

@@

Completion rate is the fundamental usability metric that measures if users can accomplish the task. We defined success as a binary metric that is true if a user connected to Tor and false if a user did not. Task time is a supplemental usability metric that measures user productivity during the task. We measured this in two ways. Connection time measures how long users took to connect to Tor, which includes the time Tor takes to establish a series of relays, while the user is forced to wait. Since time spent waiting can dominate the measurement, we use configuration time to measure the time users spent actively configuring bridges and proxies.

\subsection{Experimental Setup}
\label{sec:environments}
We simulated three environments of various censorship severity (Table~\ref{tab:environments}). We chose to 
simulate environments for the stability and reproducibility of the experiment, as real censored networks are volatile and complex.

@@

In the first environment, E1, particular websites were blocked, which meant that they could be accessed via a default Tor configuration (i.e., bridges or proxies were not needed). In the second environment, E2, we additionally blocked the IP addresses of public Tor relays, which meant that Tor could still be accessed by enabling any of the included bridges. Finally, the third environment, E3, additionally blocked several of the included bridges, which meant experimenting with several different bridges---beyond the recommended ones---in order to successfully connect.

The simulated environments are not intended to imitate any particular country's network---though they do use the same techniques that are currently used by various censorship regimes. Rather, we designed simulated environments to require different bridges to connect to Tor. We believe this to be sufficient for the purpose of testing the configuration interface, since this requires the user to take the interface paths we wanted to test (and matches the actions real users would need to take to circumvent various types of censorship).

@@

\begin{itemize}
\item {\bfseries Technical terms:} The interface uses technical and Tor-specific terms that are likely unfamiliar to an average user (e.g., ``Tor network,'' ``local proxy,'' ``bridges,'' and ``transports''). 
\item {\bfseries Too many choices:}  Users who do not need a bridge or proxy are able to configure them. There are nine choices for bridges by transport and more options to input a custom bridge. Almost any HTTP or SOCKS proxy can be used as the proxy. Custom bridge lines require the bridge information with specific syntax, but this syntax is not specified in the interface. Proxy fields also do not provide syntactic guidelines. See the whole list of possible errors in Appendix~\ref{states}. 
\item {\bfseries Poor feedback:} Even if there was a syntactic error in the bridge or proxy fields, the user is not notified until after a connection is attempted and fails. If a censored bridge was chosen, the connection timeout takes several minutes. Error messages were often technical and did not give suggestions.
\end{itemize}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{torconfig.pdf}
\caption{
A digraph illustrating the potential user paths through the interface. Non-leaf nodes are screens in the user interface (start, bridge 1, bridge 2, proxy 1, proxy 2, and progress). Leaf nodes are various end states (quit, success, or error). Transitions between these nodes denote user actions. For details of each end state, see Appendix~\ref{states}. 
}
\label{fig:digraph}
\end{figure*}

@@

We talk about the main reasons why participants struggled to use Tor launcher, summarizing general behavioral trends and attitudes. We detail what each participant did in Appendix~\ref{summary}.

\subsection{Pain Points in Tor Launcher} 
\label{sec:pain-points}
Participants spent a lot of time going back and forth on screens, looking for proxy information, and waiting for the connection to work or timeout. We believe that the reasons for these behaviors were:\\

@@

\subsection{Results} 
Users should be able to connect to Tor within a few minutes. The interface does not meet this goal (Table~\ref{table:participant-summary}). Our design changes helped more users connect to Tor in less time, but we still encourage further improvements. In this section, we use OLD to refer to the original interface (Figure~\ref{fig:old-interface}) and NEW to refer the version with our changes (Figure~\ref{fig:new-interface}).

We found that 63\% (72 of 114) of first attempts to connect failed and 79\% (363 of 458) of total attempts to connect failed. Participants averaged over 3 minutes in censored environments (E2) and over 20 minutes in heavily censored environments (E3, Figure~\ref{fig:all-participant-edges}). Based on this, we recommend a thorough redesign of the interface.

@@

For our work, we draw on the large body of previous work in secure interaction design, computer-human interaction, and user studies. Work we found useful to our experiment includes research on principles for secure interaction design~\cite{yee2002user}, common problems in human-computer dialogue design~\cite{molich1990improving}, user tolerance of security delays~\cite{egelmanplease}, treating human attention as a scarce resource when to delegating tasks to users~\cite{bohme2011security}, stopping users from doing bad things~\cite{good2007noticing}, and getting proper user consent~\cite{bohme2010trained}.

@@

\section {Acknowledgments}
Georg Koppen, Mike Perry, and Isabela Bagueros from Tor Project, Inc. provided valuable insights about the history of Tor launcher and internal constraints.

>>>>

\section{Introduction}
Tor~\cite{dingledine2004tor} is an anonymity network that routes Internet traffic through a series of relays 
to make it difficult to observe sources and destinations. Tor Browser~\cite{torbrowser} is a modified Firefox browser that routes traffic through Tor, thereby simplifying the process of configuring Tor to use existing browsers. Although Tor Browser was originally designed for anonymity, many now use Tor Browser to circumvent Internet censorship. Many countries block Tor relays specifically to prevent their citizens from circumventing censorship~\cite{winter2012great}. Tor Browser is bundled with a graphical configuration interface, Tor Launcher (Figure~\ref{fig:old-interface}), which allows users to bypass censorship by configuring various bridges, proxies, and transport protocols prior to connecting to Tor. Of course, these configuration options are only effective if users understand how to correctly use them. Thus, we performed a usability study wherein we simulated various censorship environments in order to examine the effectiveness of Tor Launcher and how it could be improved.

Improving Tor launcher has the obvious benefit of helping censored users connect to Tor, but this also benefits other Tor users since an increase in the overall number of users increases the anonymity set~\cite{dingledine2006anonymity}). In this paper, we describe how we started our usability evaluation by inspecting the interface (Section~\ref{sec:inspection}) and observing participants interact with it (Section~\ref{sec:qualitative}). Based on the insights from our observations, we proposed several changes to the interface (Section~\ref{sec:design}) and quantified how usable the interface was before and after changes (Section~\ref{sec:quantitative}). Finally, we provide recommendations based on what we learned (Section~\ref{sec:recommendations}), we address limitations of our evaluation (Section~\ref{sec:limitations}), and frame our study in the context of existing work (Section~\ref{sec:related}). We hope that our usability evaluation serves as an informative case study and provides guidance on further improving Tor Launcher, as well as other privacy-enhancing technologies.

@@

Bridges are unlisted alternative entry nodes used to connect to Tor when Tor relays are censored.
Bridges also receive packets from the user and forward traffic to middle relays, but may also run a pluggable transport (referred to as ``transport'' for shorthand). Pluggable transports specify alternative protocols by which Tor Browser may connect to the Tor network, thereby disguising Tor traffic to make it harder to detect and block bridges. For convenience, bridges are grouped by which transport they run.

Most bridges can be blocked by blocking their static IP addresses. These include bridges that run ``fte'' and ``fte-ipv6''~\cite{fte} pluggable transports, which disguise the Tor protocol as another protocol (such as HTTP); these also include the ``obfs3''~\cite{obfs3}, ``obfs4''~\cite{obfs4}, and ``scramblesuit''~\cite{scramblesuit} transports, which encrypt or alter the Tor protocol to appear as random noise. Other bridges are less susceptible to blocking because of how they work. Bridges that run the ``flashproxy''~\cite{flashproxy} and ``meek''~\cite{fifield2015blocking} transports route traffic through third party web browsers and content delivery networks. The bridges that are accessible in each country vary. Figure~\ref{fig:old-bridge-combobox} shows the bridge and transport options at the time of our study.

@@

\subsection{Connecting to Tor} 
It is important to note that there are many valid configuration settings that can be used to connect to the Tor network. A user who does not need a bridge or proxy can still connect with a bridge, proxy, or both, provided that they are configured correctly (though adding unnecessary network hops or obfuscation protocols may slow the connection down).

@@

\subsection{Testing Pipeline} 
We performed a cognitive walkthrough~\cite{wharton1994cognitive,cognitive-walkthrough} to inspect the interface to anticipate potential problems and formally assess its pros and cons. We then conducted qualitative and quantitative usability tests. We believe both methods were necessary for a complete evaluation. Qualitative testing tells you {\it why or how} it is or is not usable, while quantitative testing tells you {\it how much} it is or is not usable. For gathering qualitative data, five users is considered to maximize the cost-benefit ratio~\cite{howmanyusers}, so we recruited five participants for each condition---multiple simulated censorship environments---in our qualitative experiment. Twenty users is considered to be the minimum for yielding statistically significant measurements~\cite{howmanyusers}, so we recruited twenty participants for each condition in our quantitative experiment.

@@

Completion rate is the fundamental usability metric that measures if users can accomplish the task. We defined success as a binary metric that is true if a user connected to Tor and false if a user did not. Task time is a supplemental usability metric that measures user productivity during the task. We measured this in two ways. Connection time measures how long users took to connect to Tor, which includes the time Tor takes to establish a series of relays, while the user is forced to wait. Since time spent waiting can dominate the measurement, we use configuration time to measure the time users spent actively configuring bridges and proxies. Connection time is always greater than the configuration time, since it includes the configuration time.

\subsection{Experimental Setup}
\label{sec:environments}
We simulated three environments of varying censorship severity (Table~\ref{tab:environments}). We chose to 
simulate environments for the stability and reproducibility of the experiment, as real censored networks are volatile and complex.

@@

In the first environment, E1, particular websites were blocked, which meant that they could be accessed via a default Tor configuration (i.e., bridges or proxies were not needed). In the second environment, E2, we additionally blocked the IP addresses of public Tor relays, which meant that Tor could still be accessed by enabling any of the included bridges. Finally, the third environment, E3, additionally blocked several of the included bridges, which required experimenting with several different bridges---beyond the recommended option---in order to successfully connect.

The simulated environments are not intended to imitate any particular country's network---though they do use the same techniques that are currently used by various censorship regimes. Rather, we designed simulated environments to have different requirements to connect to Tor. We believe this to be sufficient for the purpose of testing the configuration interface, since this requires the user to take the different interface paths we wanted to test (and matches the actions real users would need to take to circumvent various types of censorship).

@@

\begin{itemize}
\item {\bfseries Technical terms:} The interface uses technical and Tor-specific terms that are likely unfamiliar to an average user (e.g., ``Tor network,'' ``local proxy,'' ``bridges,'' and ``transports''). 
\item {\bfseries Too many choices:}  Users who do not need a bridge or proxy are able to configure them. There are nine choices for bridges by transport and more options to input a custom bridge. Almost any HTTP or SOCKS proxy can be used as the proxy. Custom bridge lines require the bridge information with specific syntax, but this syntax is not specified in the interface. Proxy fields also do not provide syntactic guidelines. See the whole list of possible errors we identified in Appendix~\ref{states}. 
\item {\bfseries Poor feedback:} Even if there was a syntactic error in the bridge or proxy fields, the user is not notified until after a connection is attempted and fails. If a censored bridge was chosen, the connection timeout takes several minutes. Error messages were often technical and did not give suggestions.
\end{itemize}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{torconfig.pdf}
\caption{
A digraph illustrating the potential user paths through the interface. Non-leaf nodes are screens in the user interface (start, bridge 1, bridge 2, proxy 1, proxy 2, and progress). Leaf nodes are various end states (quit, success, or error). Transitions between these nodes denote user actions. 
}
\label{fig:digraph}
\end{figure*}

@@

Here we report about the main reasons why participants struggled to use Tor launcher, summarizing general behavioral trends and attitudes. We detail what each participant did in Appendix~\ref{summary}.

\subsection{Pain Points in Tor Launcher} 
\label{sec:pain-points}
Participants spent a lot of time going back and forth on screens, looking for proxy information, and waiting for the connection to work or time out. We believe that the reasons for these behaviors were:\\

@@

\subsection{Results} 
In this section, we use OLD to refer to the original interface (Figure~\ref{fig:old-interface}) and NEW to refer the version with our changes (Figure~\ref{fig:new-interface}). Users should be able to connect to Tor within a few minutes. The interface does not meet this goal (Table~\ref{table:participant-summary}). Our design changes helped more users connect to Tor in less time, but we still encourage further improvements.

We found that 63\% (72 of 114) of first attempts to connect failed and 79\% (363 of 458) of total attempts to connect failed. Even with the redesigned interface, participants averaged over 3 minutes in censored environments (E2) and over 20 minutes in heavily censored environments (E3, Figure~\ref{fig:all-participant-edges}). Based on this, we recommend a thorough redesign of the interface.

@@

For our work, we draw on the large body of previous work in secure interaction design, computer-human interaction, and user studies. Work we found useful to our experiment includes research on principles for secure interaction design~\cite{yee2002user}, common problems in human-computer dialogue design~\cite{molich1990improving}, user tolerance of security delays~\cite{egelmanplease}, treating human attention as a scarce resource when to delegate tasks to users~\cite{bohme2011security}, stopping users from installing potentially harmful programs~\cite{good2007noticing}, and getting proper user consent~\cite{bohme2010trained}.

@@

\section {Acknowledgments}
We would like to thank the peer reviewers, especially our shepherd Jens Grossklas, for giving us valuable feedback on this paper. Tor Project employees Georg Koppen, Mike Perry, and Isabela Bagueros provided valuable insights about the history of Tor launcher and internal constraints. This work was done in collaboration with Xlab (Experimental Social Science Laboratory)~\cite{Xlab}, SCRUB (secure computing research for users' benefit)~\cite{scrub}, and BLUES (Berkeley laboratory for usable and experimental security)~\cite{blues} laboratories, with funding from Intel and the National Science Foundation.
